---
title: "covid 19"
author: "Wilhelmus Elersich"
date: "9/27/2022"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Question 1 - Is dying independent of sex
H_0 dying of Covid-19 is independent of sex

H_1 dying of Covid-19 is dependent on sex



Table of patients:
```{r}

# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```

## We create two bar charts - one based on frequency and the other on percent, use ggplot commands to do this

```{r}
chartdata <- kRdata %>%
mutate(across(where(is.character), as.factor))
chart<- chartdata %>% group_by(result,Sex) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


chartfreq <- ggplot(chart,aes(x=result,y=count,fill=Sex))
chartfreq + geom_bar(stat="identity",position="dodge")

chartperc <- ggplot(chart,aes(x=result,y=prcnt*100,fill=Sex))
chartperc + geom_bar(stat="identity",position="dodge")

```

Here we can see that the percentage based bar chart would be more useful for visualizing our data. By looking at this chart, we can see that a greater percentage of those killed were male. It would appear that Covid-19 is more deadly to males than to females.





  
## Numerical results
```{r}
table1 <- xtabs(~result + Sex, data=chartdata)
rowPerc(table1)
colPerc(table1)
```

The lower table does not seem to be very useful to answer our question. Looking at the upper plot, we can conclude that males are dying more than females, with 64% of those dying being male.

## Inferential results 
We shall use the Chi-squared test and the fisher test.

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```
The low p-values tell us that there is a very low chance that the data we have is coincidental. The odds ratio states that males are 2.3 times more likely to die from Covid-19 than females, which is backed by the confidence interval. We can Therefore we reject the null hypothesis, and state that whether or not one dies from Covid-19 is dependent on sex. 


## Conclusions
From our graphical, numerical, and inferential results we determined that men are more than twice as likely as women to die from Covid-19.
