---
title: "covid 19"
author: "Wilhelmus Elersich"
date: "9/27/2022"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Question 1 - Is dying independent of sex




Table of patients:
```{r}

# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```

## We create two bar charts - one based on frequency and the other on percent, use ggplot commands to do this

```{r}
chartdata <- kRdata %>%
mutate(across(where(is.character), as.factor))
chart<- chartdata %>% group_by(result,Sex) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


chartfreq <- ggplot(chart,aes(x=result,y=count,fill=Sex))
chartfreq + geom_bar(stat="identity",position="dodge")

chartperc <- ggplot(chart,aes(x=result,y=prcnt*100,fill=Sex))
chartperc + geom_bar(stat="identity",position="dodge")

```

Here we can see that the percentage based bar chart would be more useful for visualizing our data. By looking at this chart, we can see that a greater percentage of those killed were male. It would appear that Covid-19 is more deadly to males than to females.





  
## Numerical results
```{r}
table1 <- xtabs(~result + Sex, data=chartdata)
rowPerc(table1)
colPerc(table1)
```

The lower table does not seem to be very useful to answer our question. Looking at the upper plot, we can conclude that males are dying more than females.

## Inferential results 




## Conclusions

